---
title: "Penguins size comparisons"
author: "Karla Garibay Garcia"
date: "11/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```



# Part 1. T-tests to compare means between two different groups. 


Asking: Do we have enough evidence to make us think that these samples are drawn from populations with different means, or am I sticking with the null hypothesis that these samples were drawn from populations with the same means?


### Mass comparison (body_mass_g): male & female chinstrap penguins

1. Explore data by:
- Making a subset that only contains chinstrap penguins
- Making basic plots to explore normality and distribution
- Get descriptive statistics


```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap")


#Histograms - don't need a y variable bc it'll automatically count the # of observations in the different bins created for body mass.Use `facet_wrap` to split the male & female groups into separate plots. 
ggplot(data = chinstraps, 
       aes(x = body_mass_g)) +
  geom_histogram(bins = 10) +
  facet_wrap(~sex)
#Histograms look somewhat normally distributed and bell shape is somewhat symmetric, so it seems that mean would be a reasonable metric of central tendency. The central tendency of the male weight is a bit higher, closer to 4,000 g , than the female central tendency, closer to 3,500 g. 


#QQplot - compares our sample to data if it were perfectly drawn from a normal distribution. The closer to linear, the closer our sample is to being normally distributed or being drawn from a normal distribution.
ggplot(data = chinstraps, 
       aes(sample = body_mass_g)) +
  geom_qq() +
  facet_wrap(~sex)
#QQplots look very close to normally distributed.


#Descriptive Statistics using group_by() and summarize, which recognize groups w/i a variable and then puts the fn that we call in summarize, condenses the calculations for each group into a single value and reports it as a nice summary table.
chinstraps %>% 
  group_by(sex) %>% 
  summarize(
    mean_mass = mean(body_mass_g), 
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g), 
    sample_size = n())
#I can see that means are a bit different for females & males, both have an equal sample size, similar sd's, median are similar to the mean which is reflective of the symmetric distribution. Looking at these differences is more important than a p-value.
 
```

2. Get vectors of body_mass_g for female & male chinstap penguins separately. 

```{r}
chinstraps_f <- chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)

chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull
```

### T-test


1. Is there  a sig diff in body mass b/w male & female chinstrap penguins? (2 sided) Do we have enough evidence to think the mean difference is far enough from 0 to think that the pop mean mass is diff for the groups that these were drawn from.


```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest #Call it to view below. The lowest line tells you the means of the two groups in the order you put them in. The null hyp is that the means are equal so the diff in means is equal to 0. The t value is a measure of how diff from 0 our means are on the t distribution when taking into account sample size and data dispersion. df are a measure of sample size. p-value is the prob that I could've gotten sample means that are at least as different as ours are by random chance if they were drawn from pops with the same mean. Since our p-value is super low, this is enough evidence to reject the null that these samples were drawn from pop with same mean body mass. Our conclusion is that there's a significant difference for mean body mass b/w male & female chinstrap penguins. It's important to add the actual difference, how much larger males are than females, and the effect size.

cohen.d(chinstraps_f, chinstraps_m) #Gives us magnitude of 1.26, which is considered large effect size. So this is a clear difference that exists between female & male chinstrap penguins. The effect size is not so highly impacted by sample size. 

#Look at key and use of broom fn

```
 
2. Are male chinstrap penguins sig larger in body mass than female chinstrap penguins? (1 sided test). What is the prob that I could've gotten a means diff that is at least this larger or greater by random chance? 


```{r}
male_ttest <- t.test(chinstraps_m, chinstraps_f, alternative = "greater") #the alternative greater or less is what fits in b/w the vectors
male_ttest #this tells me: yes, reject the null that the means difference is not greater than 0 aka it's the same, and retain the alt hyp that the means diff is greater than 0. The way to write it: males mean chinstrap penguins body mass is sig greater than female mean chinstrap body mass, and then report the significance level.
```


# Part 2. One-way ANOVA (way to compare means across >2 groups)


### Looking at penguins sizes, specifically mean flipper length of female penguins b/w the three species 


1. Explore data by:
- Make a table of counts by species and sex
- Create subset of only female penguins
- Look at distributions of each w/ histrogram and qqplot
- Summary statistics


```{r}
penguin_counts <- penguins %>% 
  count(species, sex) #list vairables that I want to create groups by. could also do this w/group_by, summarize and n.
penguin_counts #n is large enough to assume CLT applies and data is normally distributed, but still need to check


penguin_f <- penguins %>% 
  filter(sex == "female")


ggplot(data = penguin_f, aes(x = flipper_length_mm)) +
  geom_histogram(aes(fill = species))+ #Look pretty bell-shaped, pretty close to normally distributed. 
  facet_wrap(~species) #a good default setting is that all of the axes have to match for each graph, which makes it easier to compare across groups. If ever exploring the distributions w/i groups, add `,scale = "free"` 


ggplot(data = penguin_f, aes(sample = flipper_length_mm)) +
  geom_qq(aes(color = species))+ #Very close to linear, so this shows my samples were drawn from populations with normal distributions. 
  facet_wrap(~species) 

penguin_f_summary <-  penguin_f %>% 
  group_by(species) %>% 
  summarize(mean_flipper_length = mean(flipper_length_mm), 
            sd_flipper_length = sd(flipper_length_mm),
            sample_size = n(), 
            se_flipper_length = sd(flipper_length_mm)/sqrt(n()),
            variance_flipper_length = var(flipper_length_mm)) #If my largest variance for a group is <4x greater than the smallest variance, then they're close enough to run an ANOVA
penguin_f_summary

leveneTest(flipper_length_mm ~ species, data = penguin_f) #Hyp test, where null is that our groups have equal variances. Looking at variance in flipper length as a fn of species. Looking at the Pr(>F) value is the same as the p-value, which here is >.05 sig so we retain the null hyp that variance is equal. 
```

2. Beeswarm plot (way to show summary statistics) w/mean +- 1sd


```{r}
ggplot() +
  geom_beeswarm(data = penguin_f, 
                aes(x = species, 
                    y = flipper_length_mm), #shows me distributions of my observations and arranges them where the greater densities of observations
                color = "cyan", 
                alpha = 0.5) +#if customizations refer to a variable it should be in aes, if not it should be outside of aes.
  geom_point(data = penguin_f_summary, #adds points on top of beeswarm plot
             aes(x = species, 
                 y = mean_flipper_length)) +
  geom_errorbar(data = penguin_f_summary, #adds whiskers
                aes(x = species, 
                    ymax = mean_flipper_length + sd_flipper_length, 
                    ymin = mean_flipper_length - sd_flipper_length), 
                width = 0.1) + #adjusts width of error bar 
  theme_minimal()
  
```

### One-way ANOVA 

Answers: Do we have enough evidence to reject the null that all means are equal? 

```{r}
penguin_aov <- aov(flipper_length_mm ~ species, 
                   data = penguin_f)
summary(penguin_aov) # F statistics is the ratio B/w group variance/ w/i group variance. For this example, the F statistic is high, which means the Bss is higher than the Wss and that's reflected in the p-value, which is a really low probability that we could've gotten sample means by random chance that are at least this different if all of our samples were drawn from populations with the same mean.So we have enough evidence to reject the null hyp that the means for all of these species are the same. B/c of this result we want to do a post-hoc test

TukeyHSD(penguin_aov)#does pairwise comparisons and adjusts so we don't have the 5% error each time. We can see that all p values are below 0.05 so we conclude there's a sig diff in female flipper length b/w all of the groups aka b/w each species.
```

## Type III ANOVA with car package 


If we have very unbalanced design, with different n and distributions w/i groups, we want to do an orthogonal or type 3 anova by: 
- Creating a linear model that explores flipper length as a fn of species


```{r}
penguin_lm <- lm(flipper_length_mm ~ species, 
                 data = penguin_f)
penguin_aov_t3 <- Anova(penguin_lm, type = "III") #tell it what type of anova you want to performk
penguin_aov_t3 #most of the time, it's similar to the one-way and type II anova, unless data is VERY unbalanced
```

Conclusion: there is enough evidence to reject the null that our samples are drawn from pops with the same mean. aka there's a sig diff in mean flipper length b/w female gentoo, chinstrap, and adelie penguins.

