---
title: "Penguins size comparisons"
author: "Karla Garibay Garcia"
date: "11/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

# Part 1. T-tests to compare means between two different groups. 


Asking: Do we have enough evidence to make us think that these samples are drawn from populations with different means, or am I sticking with the null hypothesis that these samples were drawn from populations with the same means?


Mass comparison (body_mass_g): male & female chinstrap penguins

1. Explore data by:
- Making a subset that only contains chinstrap penguins
- Making basic plots to explore normality and distribution

- Get descriptive statistics


```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap")


#Histograms - don't need a y variable bc it'll automatically count the # of observations in the different bins created for body mass.Use `facet_wrap` to split the male & female groups into separate plots. 
ggplot(data = chinstraps, 
       aes(x = body_mass_g)) +
  geom_histogram(bins = 10) +
  facet_wrap(~sex)
#Histograms look somewhat normally distributed and bell shape is somewhat symmetric, so it seems that mean would be a reasonable metric of central tendency. The central tendency of the male weight is a bit higher, closer to 4,000 g , than the female central tendency, closer to 3,500 g. 


#QQplot - compares our sample to data if it were perfectly drawn from a normal distribution. The closer to linear, the closer our sample is to being normally distributed or being drawn from a normal distribution.
ggplot(data = chinstraps, 
       aes(sample = body_mass_g)) +
  geom_qq() +
  facet_wrap(~sex)
#QQplots look very close to normally distributed.


#Descriptive Statistics using group_by() and summarize, which recognize groups w/i a variable and then puts the fn that we call in summarize, condenses the calculations for each group into a single value and reports it as a nice summary table.
chinstraps %>% 
  group_by(sex) %>% 
  summarize(
    mean_mass = mean(body_mass_g), 
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g), 
    sample_size = n())
#I can see that means are a bit different for females & males, both have an equal sample size, similar sd's, median are similar to the mean which is reflective of the symmetric distribution. Looking at these differences is more important than a p-value.
 
```



